---
title: "W_UP_01"
author: "Josephh Diaz"
date: "fall 2025"
output: html_document
---




The report structure should include five sections.  see below



```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
mydata <- read.csv(file="pushups.csv")
```

# Introduction

<!--In this section you explain what you are trying to show. Where did the data come from? What is the research or other question you are trying to answer?!-->

Create an Rstudio project in your shared space on rstudio.cloud.  Using the dataset called pushups from the tigerstats package in R; answer the question posed by this null and alternate hypothesis in an html markdown document.

$H_0$: There is no difference between the average number of pushups of line vs skill players

$H_A$: The average number of pushups done by line players is statistically different than that of skill players.

# Methods

<!--Decide on your methods: use "variable analysis" or other appropriate descriptors. Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections: One for your descriptive results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here. Make sure to show your code. Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too. Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(pushups ~ position,
        data = mydata,
        main = "Pushups by Player Type",
        xlab = "Player Type",
        ylab = "Number of Pushups",
        col = c("lightblue", "lightgreen"))

```


### Numerical Descriptive Results


```{r}
mydata %>%
  group_by(position) %>%
  summarise(
    n = n(),
    mean = mean(pushups),
    sd = sd(pushups),
    min = min(pushups),
    max = max(pushups)
  )



```
Line players completed an average of 17.1 pushups (SD = 9.7, range 6–34, n = 8).
Skill players completed an average of 49.7 pushups (SD = 23.0, range 18–108, n = 22).
These descriptive results show that skill players tend to complete much more pushups than line players, and they also show greater variation in performance.


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results



```{r}

t_result <- t.test(pushups ~ position, data = mydata)
t_result


```
A Welch two-sample t-test was conducted to compare the mean number of pushups completed by line players and skill players. The test produced a t-statistic of t = 5.43 with approximately 10.6 degrees of freedom and a p-value of p = 0.00000942.

This p-value represents the probability of observing a difference in sample means as large as the one we found, or larger, if the null hypothesis were actually true—that is, if line and skill players truly had the same average number of pushups in the population.

In other words, if there were no real difference, the chance of obtaining a difference this extreme from random sampling alone is about 0.00094% (less than one-thousandth of one percent).

Because this probability is extraordinarily small, we reject the null hypothesis and conclude that skill players and line players do not perform the same average number of pushups.



<!--State hypothesis clearly. Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc. Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

The descriptive results showed a clear difference between the two groups, with skill players completing substantially more pushups on average than line players. The inferential test confirmed that this observed difference is extremely unlikely to have occurred by random chance alone if the population means were truly equal.

Based on the very small p-value from the Welch t-test, we reject the null hypothesis and conclude that there is strong statistical evidence of a real difference in average pushup performance between line and skill players.

This study is observational, so we cannot conclude that being a skill player causes higher pushup totals—other factors such as training, conditioning, or positional demands may explain the difference. However, the evidence strongly suggests that the two groups differ in their typical physical performance as measured by pushups.

<!--Discussion and conclusion here. If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason. Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

